Effectiveness of visual test in linear regression diagnositics

### 2022-07-30 paper_comparison/paper_comparison/paper_comparison.Rmd
modify power plot

@@

With our knowledge, what haven't been examined so far is the effectiveness of visual test relative to the equivalent conventional test in regression diagnostics. Particularly, its ability to detect non-linearity and heteroskedasticity compared to F-test and BP-test. For this purpose, two experiments were conducted. The first experiment has ideal scenario for conventional testing, where we would not expect the visual test to do better than the conventional test. The second experiment is a scenario where the conventional test is an approximate test, which visual test may have a chance to match the performance of the conventional test.

@@

For each parameter combination, effect $E$ is derived from the Kullback-Leibler divergence (the derivation is provided in [appendix ref here]) formulated as

@@

The logistic regression is fit using $log_e(E)$ as the only fixed effect covariate for the power of visual test formulated as

@@

Table [table ref here] shows the parameter estimates of the logistic regression. [Discussion about the numeric estimates here]

@@

Figure [figure ref here] shows an overview of estimated power against natural logarithm

```{r}
conv_test_plot <- function() {
  conv_result %>%
  ggplot() +
  stat_smooth(geom = "line", 
              aes(log(effect_size), reject, group = boot_id, col = "F-test"), 
              se = FALSE, method = "glm", method.args = list(family = binomial()), alpha = 0.05) +
  stat_smooth(geom = "line", 
              aes(log(effect_size), reject, col = "F-test"), 
              se = FALSE, method = "glm", method.args = list(family = binomial()), alpha = 1) +
  labs(col = "Test") +
  theme_light() +
  xlab(quote(log[e](E))) +
  ylab("Power")
}

conv_test_plot()
```

```{r}
add_wrong_BP_conv_test_plot_layers <- function(p) {
  p +
  stat_smooth(data = wrong_BP_conv_result, geom = "line", 
              aes(log(effect_size), reject, group = boot_id, col = "BP-test"), 
              se = FALSE, method = "glm", method.args = list(family = binomial()), alpha = 0.05) +
  stat_smooth(data = wrong_BP_conv_result, geom = "line", 
              aes(log(effect_size), reject, col = "BP-test"), 
              se = FALSE, method = "glm", method.args = list(family = binomial()), alpha = 1)
}

conv_test_plot() %>%
  add_wrong_BP_conv_test_plot_layers()
```

```{r}
add_wrong_WS_conv_test_plot_layers <- function(p) {
  p +
  stat_smooth(data = wrong_WS_conv_result, geom = "line", 
              aes(log(effect_size), reject, group = boot_id, col = "WS-test"), 
              se = FALSE, method = "glm", method.args = list(family = binomial()), alpha = 0.05) +
  stat_smooth(data = wrong_WS_conv_result, geom = "line", 
              aes(log(effect_size), reject, col = "WS-test"), 
              se = FALSE, method = "glm", method.args = list(family = binomial()), alpha = 1)
}

conv_test_plot() %>%
  add_wrong_BP_conv_test_plot_layers() %>%
  add_wrong_WS_conv_test_plot_layers()
```

@@

add_visual_test_plot_layers <- function(p) {
  visual_dat <- polynomials %>%
  group_by(lineup_id) %>%
  summarise(across(c(type:p_value), first)) %>%
  filter(lineup_id < 577) %>%
  mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>%
  boot_dat(times = 100)

  p +
  stat_smooth(data = filter(visual_dat, boot_id == 1), geom = "line", 
              aes(log(effect_size), reject, group = boot_id, col = "Visual"), 
              method = "glm", method.args = list(family = binomial), se = FALSE, alpha = 1) +
  stat_smooth(data = visual_dat, geom = "line", 
              aes(log(effect_size), reject, group = boot_id, col = "Visual"), 
              method = "glm", method.args = list(family = binomial), se = FALSE, alpha = 0.05)

@@

control_col <- function(p, ...) {
  p + scale_color_manual(values = unlist(list(...)))
}

conv_test_plot() %>%
  add_wrong_BP_conv_test_plot_layers() %>%
  add_wrong_WS_conv_test_plot_layers() %>%
  add_visual_test_plot_layers()
```

>>>>

## Effectiveness of visual test in regression diagnostics

@@

With our knowledge, what haven't been examined so far is the effectiveness of visual test relative to the equivalent conventional test in regression diagnostics. Particularly, its ability to detect non-linearity and heteroskedasticity compared to F-test and BP-test.

For the purpose of examining the effectiveness of visual test in regression diagnostics, two experiments were conducted. The first experiment has ideal scenario for conventional testing, where we would not expect the visual test to do better than the conventional test. The second experiment is a scenario where the conventional test is an approximate test, which visual test may have a chance to match the performance of the conventional test.

@@

For each parameter combination, effect $E$ is derived from the Kullback-Leibler divergence (the derivation is provided in [appendix ref here]) formulated as:

@@

The logistic regression is fit using $log_e(E)$ as the only fixed effect covariate for the power of visual test formulated as:

@@

The same model is fit on different subsets of the collated data to study the effect of different factors have on the power of the visual test.

Table [table ref here] shows the parameter estimates of the logistic regressions. [Discussion about the numeric estimates here]

@@

Figure \ref{fig:power-overview} shows an overview of estimated power of visual test against natural logarithm of the effect. Without further concenring different component contributing to the effect,

```{r}
conv_test_plot <- function() {
  conv_result %>%
  ggplot() +
  stat_smooth(geom = "line", 
              aes(log(effect_size), reject, group = boot_id, col = "F-test"), 
              se = FALSE, method = "glm", method.args = list(family = binomial()), alpha = 0.05) +
  stat_smooth(geom = "line", 
              aes(log(effect_size), reject, col = "F-test"), 
              se = FALSE, method = "glm", method.args = list(family = binomial()), alpha = 1) +
  labs(col = "Test") +
  theme_light(base_size = 5) +
  xlab(quote(log[e](E))) +
  ylab("Power")
}

add_wrong_BP_conv_test_plot_layers <- function(p) {
  p +
  stat_smooth(data = wrong_BP_conv_result, geom = "line", 
              aes(log(effect_size), reject, group = boot_id, col = "BP-test"), 
              se = FALSE, method = "glm", method.args = list(family = binomial()), alpha = 0.05) +
  stat_smooth(data = wrong_BP_conv_result, geom = "line", 
              aes(log(effect_size), reject, col = "BP-test"), 
              se = FALSE, method = "glm", method.args = list(family = binomial()), alpha = 1)
}

add_wrong_WS_conv_test_plot_layers <- function(p) {
  p +
  stat_smooth(data = wrong_WS_conv_result, geom = "line", 
              aes(log(effect_size), reject, group = boot_id, col = "WS-test"), 
              se = FALSE, method = "glm", method.args = list(family = binomial()), alpha = 0.05) +
  stat_smooth(data = wrong_WS_conv_result, geom = "line", 
              aes(log(effect_size), reject, col = "WS-test"), 
              se = FALSE, method = "glm", method.args = list(family = binomial()), alpha = 1)
}
```

@@

add_visual_test_plot_layers <- function(p) {
  visual_dat <- polynomials %>%
  group_by(lineup_id) %>%
  summarise(across(c(type:p_value, prop_detect), first)) %>%
  filter(lineup_id < 577) %>%
  mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>%
  boot_dat(times = 100)

  p +
  stat_smooth(data = filter(visual_dat, boot_id == 1), geom = "line", 
              aes(log(effect_size), reject, group = boot_id, col = "Visual"), 
              method = "glm", method.args = list(family = binomial), se = FALSE, alpha = 1) +
  stat_smooth(data = visual_dat, geom = "line", 
              aes(log(effect_size), reject, group = boot_id, col = "Visual"), 
              method = "glm", method.args = list(family = binomial), se = FALSE, alpha = 0.05) +
  geom_point(data = filter(visual_dat, boot_id == 1), aes(log(effect_size), prop_detect), alpha = 0.1)

@@

control_col <- function(p, ...) {
  p + scale_color_manual(values = unlist(list(...)))
}
```

```{r eval = FALSE}
polynomials %>%
  group_by(lineup_id) %>%
  summarise(dd = first(abs(effect_size - exp(2)))) %>%
  arrange(dd)
```

```{r power-overview, fig.height=7.5, fig.cap="Power overview. \\label{fig:power-overview}"}
conv_test_plot() %>%
  add_wrong_BP_conv_test_plot_layers() %>%
  add_wrong_WS_conv_test_plot_layers() %>%
  add_visual_test_plot_layers() +
  geom_vline(xintercept = 0) + 
  geom_vline(xintercept = 2) +
  geom_vline(xintercept = 4) +
  geom_vline(xintercept = 6) -> base_plot

VI_MODEL$plot_lineup(polynomials_lineup$lineup_314$data,
                     size = 0.1,
                     stroke = 0.2,
                     remove_grid_line = TRUE,
                     theme = theme_light(base_size = 5),
                     remove_axis = TRUE) +
  ggtitle(bquote(log[e](E) == 0~", F-test p-value" == .(format(filter(polynomials, lineup_id == 314)$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(polynomials, lineup_id == 314)$p_value[1], digits = 3)))) -> bottom_plot_1

VI_MODEL$plot_lineup(polynomials_lineup$lineup_415$data,
                     size = 0.1,
                     stroke = 0.2,
                     remove_grid_line = TRUE,
                     theme = theme_light(base_size = 5),
                     remove_axis = TRUE) +
  ggtitle(bquote(log[e](E) == 2~","~"F-test p-value" == .(format(filter(polynomials, lineup_id == 415)$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(polynomials, lineup_id == 415)$p_value[1], digits = 3)))) -> bottom_plot_2

VI_MODEL$plot_lineup(polynomials_lineup$lineup_494$data,
                     size = 0.1,
                     stroke = 0.2,
                     remove_grid_line = TRUE,
                     theme = theme_light(base_size = 5),
                     remove_axis = TRUE) +
  ggtitle(bquote(log[e](E) == 4~","~"F-test p-value" == .(format(filter(polynomials, lineup_id == 494)$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(polynomials, lineup_id == 494)$p_value[1], digits = 3)))) -> bottom_plot_3

VI_MODEL$plot_lineup(polynomials_lineup$lineup_101$data,
                     size = 0.1,
                     stroke = 0.2,
                     remove_grid_line = TRUE,
                     theme = theme_light(base_size = 5),
                     remove_axis = TRUE) +
  ggtitle(bquote(log[e](E) == 6~","~"F-test p-value" == .(format(filter(polynomials, lineup_id == 101)$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(polynomials, lineup_id == 101)$p_value[1], digits = 3)))) -> bottom_plot_4

design <- c(patchwork::area(1, 1, 1, 8), 
            patchwork::area(2, 1, 2, 4), 
            patchwork::area(2, 5, 2, 8), 
            patchwork::area(3, 1, 3, 4), 
            patchwork::area(3, 5, 3, 8))

patchwork::wrap_plots(base_plot, B = bottom_plot_1, C = bottom_plot_2, D = bottom_plot_3, E = bottom_plot_4, design = design)
```

### 2022-07-30 paper_comparison/paper_comparison/paper_comparison.tex
modify power plot

@@

With our knowledge, what haven't been examined so far is the
effectiveness of visual test relative to the equivalent conventional
test in regression diagnostics. Particularly, its ability to detect
non-linearity and heteroskedasticity compared to F-test and BP-test. For
this purpose, two experiments were conducted. The first experiment has
ideal scenario for conventional testing, where we would not expect the
visual test to do better than the conventional test. The second
experiment is a scenario where the conventional test is an approximate
test, which visual test may have a chance to match the performance of
the conventional test.

@@

For each parameter combination, effect \(E\) is derived from the
Kullback-Leibler divergence (the derivation is provided in {[}appendix
ref here{]}) formulated as

@@

The logistic regression is fit using \(log_e(E)\) as the only fixed
effect covariate for the power of visual test formulated as

@@

Table {[}table ref here{]} shows the parameter estimates of the logistic
regression. {[}Discussion about the numeric estimates here{]}

@@

Figure {[}figure ref here{]} shows an overview of estimated power
against natural logarithm

@@

\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-4-1.pdf}

>>>>

\hypertarget{effectiveness-of-visual-test-in-regression-diagnostics}{%
\subsection{Effectiveness of visual test in regression
diagnostics}\label{effectiveness-of-visual-test-in-regression-diagnostics}}

With our knowledge, what haven't been examined so far is the
effectiveness of visual test relative to the equivalent conventional
test in regression diagnostics. Particularly, its ability to detect
non-linearity and heteroskedasticity compared to F-test and BP-test.

@@

For the purpose of examining the effectiveness of visual test in
regression diagnostics, two experiments were conducted. The first
experiment has ideal scenario for conventional testing, where we would
not expect the visual test to do better than the conventional test. The
second experiment is a scenario where the conventional test is an
approximate test, which visual test may have a chance to match the
performance of the conventional test.

@@

For each parameter combination, effect \(E\) is derived from the
Kullback-Leibler divergence (the derivation is provided in {[}appendix
ref here{]}) formulated as:

@@

The logistic regression is fit using \(log_e(E)\) as the only fixed
effect covariate for the power of visual test formulated as:

@@

The same model is fit on different subsets of the collated data to study
the effect of different factors have on the power of the visual test.

Table {[}table ref here{]} shows the parameter estimates of the logistic
regressions. {[}Discussion about the numeric estimates here{]}

@@

Figure \ref{fig:power-overview} shows an overview of estimated power of
visual test against natural logarithm of the effect. Without further
concenring different component contributing to the effect,

\begin{figure}
\centering
\includegraphics{paper_comparison_files/figure-latex/power-overview-1.pdf}
\caption{Power overview. \label{fig:power-overview}}
\end{figure}

\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-9-1.pdf}

@@

\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-10-1.pdf}
